Comparing software variability across and within fMRI analysis packages

### 2022-01-03 paper/paper.tex
comment methods

As a result of these factors, the variability resulting from the use of
different fMRI analysis tools implementing similar analytical approaches
(between-tool software variability), or different versions of the same tool,
 can reach worrying magnitudes. For instance, the
study in~\cite{bowring2019exploring} compared the results produced by the
three main fMRI analysis toolboxes, namely SPM~\cite{penny2011statistical},
AFNI~\cite{cox1996afni} and FSL~\cite{jenkinson2012fsl}, using similar
pipelines. It reported limited similarity between the activation clusters
produced by these tools, measured by Dice coefficients ranging from 0.0 to
0.769, where 0 indicates non-overlapping
clusters and 1 means identical clusters. More recently, the work
in~\cite{Li2021.12.01.470790} also showed a low similarity between the
results produced by five different tools (ABCD~\cite{feczko2021adolescent},
CCS~\cite{xu2015connectome}, CPAC~\cite{craddock2013towards}, DPARSF~\cite{yan2010dparsf},
fMRIPrep~\cite{esteban2019fmriprep}) and identified the main
factors contributing to these differences. The magnitude of the highlighted
differences suggest that between-tool software variability may be playing a critical
role in the reproducibility of fMRI analysis overall.

@@

%%%%%%%%%% Summary of statstics %%%%%%%%
\setlength{\tabcolsep}{4pt}
\begin{table}[h]
  \centering
  \begin{tabular}{|c|l|c|c|c|}
    \hline
    %        \multirow{2}{*}{} & \multicolumn{1}{c}{Thresholded}& & \multicolumn{1}{c}{Unthresholded}& \\
    \multicolumn{2}{|c|}{} & FSL                                & AFNI       & SPM                     \\
    \hline
    {Preprocessing}        & {Motion Correction}                & \checkmark & \checkmark & \checkmark \\
    {}                     & {Segmentation}                     &            &            & \checkmark \\
    {}                     & {Brain Extraction (Anatomical)}    & \checkmark & \checkmark & \checkmark \\
    {}                     & {Brain Extraction (Functional)}    &            & \checkmark &            \\
    {}                     & {Intra-subject Coregistration}     & \checkmark & \checkmark & \checkmark \\
    {}                     & {Inter-subject Registration}       & \checkmark & \checkmark & \checkmark \\
    {}                     & {Analysis Voxel Size}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Smoothing}                        & \checkmark & \checkmark & \checkmark \\
    \hline
    {First-level}          & {Model Specification}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Inclusion of 6 Motion Parameters} & \checkmark & \checkmark & \checkmark \\
    {}                     & {Model Estimation}                 & \checkmark & \checkmark & \checkmark \\
    {}                     & {Contrasts}                        & \checkmark & \checkmark & \checkmark \\
    \hline
    {Second-level}         & {Model Specification}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Model Estimation}                 & \checkmark & \checkmark & \checkmark \\
    {}                     & {Contrasts}                        & \checkmark & \checkmark & \checkmark \\
    {}                     & {Second-level Inference}           & \checkmark & \checkmark & \checkmark \\
    \hline
  \end{tabular}
  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}).}
  \label{table:pipeline-steps}
\end{table}

@@

We measured between-tool software variability (BT) by running the pipelines
described in~\cite{bowring2019exploring} with FSL version 5.0.10, AFNI
version 18.1.09, and SPM12 version r7771 executed with GNU/Octave version
5.2. These software versions were identical to the ones used 
in~\cite{bowring2019exploring} except for SPM for which we used GNU/Octave
instead of MATLAB to enable mathematical function instrumentation using
Fuzzy Libmath. Indeed, MATLAB uses its own built-in mathematical functions,
which prevents the use of Fuzzy Libmath. In AFNI, we set the number of
threads to 7 even though AFNI executions
in~\cite{bowring2019exploring} were single-threaded. This was meant to
reduce the time overhead resulting from Fuzzy Libmath instrumentation.
% \gk{Did we verify that single-threading generally didn't lead to different
% results than 7-threading? As in, were they all ostensibly sampled from the
% same random distribution?} \ali{We verified that multithreading leads to
% different results (e.g. single-threading vs, 7-threading). However,
% single-threaded executions couldn't reproduce the original results.}
% \tristan{Fig S1 answers this comment and is refered later on}
All the analyses were conducted on the CentOS 7.3 operating system. The
computations were performed on \href{https://www.computecanada.ca}{Compute
  Canada's} Béluga cluster nodes, each with 2$\times$ Intel Gold 6148 Skylake~@~2.4~GHz
(40 cores/node) CPU and 8~GB of RAM per core. To facilitate portability and reproducibility,
we encapsulated the
above-mentioned software packages in Docker container images based on CentOS 7.3
which we converted to Singularity images to enable running on cluster nodes.

We simulated within-tool software variability (WT) by running the same analyses three
times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
double-precision values and $t=24$~bits for single-precision values. These
values were chosen such that the numerical perturbation simulates machine
error. The resulting samples are equally plausible estimates of
the true numerical result at the precision used by the pipelines. 
% \yohan{How can you assert that when you don't have ground truth? Would it be more safe to say "plausible estimates numerical solutions"?}
% \tristan{This does not refer to the ground truth but to the true \emph{numerical} result}
% Moreover, to evaluate numerical perturbations at different magnitudes,
% we repeated the FSL analyses for virtual
% precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
% and double-precision values. We also repeated the FSL analyses for virtual
% precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having
% set the virtual precision to $t=24$~bits for single-precision values.

>>>>

As a result of these factors, the variability resulting from the use of
different fMRI analysis tools implementing similar analytical approaches
(between-tool software variability), or different versions of the same tool,
 can reach worrying magnitudes. For instance, the
study in~\cite{bowring2019exploring} compared the results produced by the
three main fMRI analysis toolboxes, namely SPM~\cite{penny2011statistical},
AFNI~\cite{cox1996afni} and FSL~\cite{jenkinson2012fsl}, using similar
pipelines. It reported limited similarity between the activation clusters
produced by these tools, measured by Dice coefficients ranging from 0.0 to
0.769, where 0 indicates non-overlapping
clusters and 1 means identical clusters. More recently, the work
in~\cite{Li2021.12.01.470790} also showed a low similarity between the
results produced by five different tools (ABCD~\cite{feczko2021adolescent},
CCS~\cite{xu2015connectome}, CPAC~\cite{craddock2013towards}, DPARSF~\cite{yan2010dparsf},
fMRIPrep~\cite{esteban2019fmriprep}) and identified the main
factors contributing to these differences. The magnitude of the highlighted
differences suggest that between-tool software variability may be playing a critical
role in the reproducibility of \camille{add ``an'' ?} fMRI analysis overall.

@@

%%%%%%%%%% Summary of statstics %%%%%%%%
\setlength{\tabcolsep}{4pt}
\begin{table}[h]
  \centering
  \begin{tabular}{|c|l|c|c|c|}
    \hline
    %        \multirow{2}{*}{} & \multicolumn{1}{c}{Thresholded}& & \multicolumn{1}{c}{Unthresholded}& \\
    \multicolumn{2}{|c|}{} & FSL                                & AFNI       & SPM                     \\
    \hline
    {Preprocessing}        & {Motion Correction}                & \checkmark & \checkmark & \checkmark \\
    {}                     & {Segmentation}                     &            &            & \checkmark \\
    {}                     & {Brain Extraction (Anatomical)}    & \checkmark & \checkmark & \checkmark \\
    {}                     & {Brain Extraction (Functional)}    &            & \checkmark &            \\
    {}                     & {Intra-subject Coregistration}     & \checkmark & \checkmark & \checkmark \\
    {}                     & {Inter-subject Registration}       & \checkmark & \checkmark & \checkmark \\
    {}                     & {Analysis Voxel Size}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Smoothing}                        & \checkmark & \checkmark & \checkmark \\
    \hline
    {First-level}          & {Model Specification}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Inclusion of 6 Motion Parameters} & \checkmark & \checkmark & \checkmark \\
    {}                     & {Model Estimation}                 & \checkmark & \checkmark & \checkmark \\
    {}                     & {Contrasts}                        & \checkmark & \checkmark & \checkmark \\
    \hline
    {Second-level}         & {Model Specification}              & \checkmark & \checkmark & \checkmark \\
    {}                     & {Model Estimation}                 & \checkmark & \checkmark & \checkmark \\
    {}                     & {Contrasts}                        & \checkmark & \checkmark & \checkmark \\
    {}                     & {Second-level Inference}           & \checkmark & \checkmark & \checkmark \\
    \hline
  \end{tabular}
  \caption{Software processing steps (adapted from~\cite{bowring2019exploring}). \camille{``Analysis Voxel Size'' is not a processing step per se but more a parametrization of inter-subject registration so maybe it should be removed? // Not sure if this is worth noting but in FSL ``Inter-subject registration'' happens after the first-level analysis}}
  \label{table:pipeline-steps}
\end{table}

@@

We measured between-tool software variability (BT) by running the pipelines
described in~\cite{bowring2019exploring} with FSL version 5.0.10, AFNI
version 18.1.09, and SPM12 version r7771 executed with GNU/Octave version
5.2. These software versions were identical to the ones used 
in~\cite{bowring2019exploring} except for SPM for which we \camille{add ``had to use a more recent version in order to be able to use it with''} GNU/Octave
instead of MATLAB to enable mathematical function instrumentation using
Fuzzy Libmath. Indeed, MATLAB uses its own built-in mathematical functions,
which prevents the use of Fuzzy Libmath. In AFNI, we set the number of
threads to 7 even though AFNI executions
in~\cite{bowring2019exploring} were single-threaded. This was meant to
reduce the time overhead resulting from Fuzzy Libmath instrumentation.
% \gk{Did we verify that single-threading generally didn't lead to different
% results than 7-threading? As in, were they all ostensibly sampled from the
% same random distribution?} \ali{We verified that multithreading leads to
% different results (e.g. single-threading vs, 7-threading). However,
% single-threaded executions couldn't reproduce the original results.}
% \tristan{Fig S1 answers this comment and is refered later on}
All the analyses were conducted on the CentOS 7.3 operating system. The
computations were performed on \href{https://www.computecanada.ca}{Compute
  Canada's} Béluga cluster nodes, each with 2$\times$ Intel Gold 6148 Skylake~@~2.4~GHz
(40 cores/node) CPU and 8~GB of RAM per core. To facilitate portability and reproducibility,
we encapsulated the
above-mentioned software packages in Docker container images based on CentOS 7.3
which we converted to Singularity images to enable running on cluster nodes.

We simulated within-tool software variability (WT) by running the same analyses three
times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
double-precision values and $t=24$~bits for single-precision values. These
values were chosen such that the numerical perturbation simulates machine
error \camille{Given the target journal where any readers (like myself :) ) won't be expert in nuemrical perturbations, I think that you could add a bit more details here on why those numbers are good to simulate machine errors}. The resulting samples are equally plausible estimates of
the true numerical result at the precision used by the pipelines. 
% \yohan{How can you assert that when you don't have ground truth? Would it be more safe to say "plausible estimates numerical solutions"?}
% \tristan{This does not refer to the ground truth but to the true \emph{numerical} result}
% Moreover, to evaluate numerical perturbations at different magnitudes,
% we repeated the FSL analyses for virtual
% precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
% and double-precision values. We also repeated the FSL analyses for virtual
% precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having
% set the virtual precision to $t=24$~bits for single-precision values.
